Railway: Adaptive Disk Storage for Interaction Graphs

### 2014-09-11 docs/edbt2015/text/introduction.tex
railroad -> railway

There a number of challenges in achieving an effective adaptive layout. First,
we need to find the partitioning of attributes that minimized the query I/O.
To address this, we model the problems of overlapping and non-overlapping
attribute partitioning as mixed-integer linear programs (ILPs), and provide
optimal solutions that minimize the query I/O cost. Second, the query
workload, and thus the attribute access pattern can change over time. For this
purpose, our railway layout supports customization of the attribute
partitioning of sub-blocks on a per-block basis. Third, such flexibility
necessitates online configuration of attribute partitioning as the query
workload evolves, which in turn requires fast algorithms for performing the
attribute partitioning. For this purpose, we develop greedy heuristic
algorithms for both overlapping and non-overlapping partitioning scenarios.
These algorithms can scale to larger number of attributes, yet provide close
to optimal query I/O performance. Finally, the railroad layout trades off
storage space to gain improved query I/O performance. The storage overhead is
more pronounced for the case of overlapping partitioning. To address this, we
limit the amount of storage  overhead that can be tolerated, and integrate
this limit to both our ILP formulations, as well as our greedy heuristics.

In summary, this paper makes the following contributions:
\begin{itemize}
\item We introduce the railroad layout for adaptive organization of interaction
graphs on disk. 
\item We introduce optimal ILP formulations for partitioning disk blocks into 
sub-blocks with overlapping and non-overlapping attributes, given a query
workload. Our formulation also support upper bounding the amount of storage
overhead introduced as a result of the railroad layout.
\item To support online adaptation, we develop greedy heuristics that can scale
better compared to the ILP alternatives, yet achieve close to optimal query
I/O.
\item We provide an extensive experimental study comparing our approach to a
few baseline alternatives.
\end{itemize}

>>>>

There a number of challenges in achieving an effective adaptive layout. First,
we need to find the partitioning of attributes that minimized the query I/O.
To address this, we model the problems of overlapping and non-overlapping
attribute partitioning as mixed-integer linear programs (ILPs), and provide
optimal solutions that minimize the query I/O cost. Second, the query
workload, and thus the attribute access pattern can change over time. For this
purpose, our railway layout supports customization of the attribute
partitioning of sub-blocks on a per-block basis. Third, such flexibility
necessitates online configuration of attribute partitioning as the query
workload evolves, which in turn requires fast algorithms for performing the
attribute partitioning. For this purpose, we develop greedy heuristic
algorithms for both overlapping and non-overlapping partitioning scenarios.
These algorithms can scale to larger number of attributes, yet provide close
to optimal query I/O performance. Finally, the railway layout trades off
storage space to gain improved query I/O performance. The storage overhead is
more pronounced for the case of overlapping partitioning. To address this, we
limit the amount of storage  overhead that can be tolerated, and integrate
this limit to both our ILP formulations, as well as our greedy heuristics.

In summary, this paper makes the following contributions:
\begin{itemize}
\item We introduce the railway layout for adaptive organization of interaction
graphs on disk. 
\item We introduce optimal ILP formulations for partitioning disk blocks into 
sub-blocks with overlapping and non-overlapping attributes, given a query
workload. Our formulation also support upper bounding the amount of storage
overhead introduced as a result of the railway layout.
\item To support online adaptation, we develop greedy heuristics that can scale
better compared to the ILP alternatives, yet achieve close to optimal query
I/O.
\item We provide an extensive experimental study comparing our approach to a
few baseline alternatives.
\end{itemize}
